---
title: "CellRanger QC: `r params$analysed_dataset`"
date: "Analysed on `r format(Sys.time(), '%B %d, %Y')`"
author: "QuantBio `r params$more_authors`"
output:
  rmdformats::downcute:
    self_contained: true
    thumbnails: false
    lightbox: true
    gallery: false
    highlight: tango
    code_folding: hide
    fig_width: 10
    fig_height: 10
    toc_depth: 5
params:
  analysed_dataset: "Example"
  more_authors: ""
  crc_mtx_dir: "./../rawData"
  qcs_patient_id: "SampleID"
  raw_seurat: "./../processedData/Su_7142_210727B6_raw.rds"
    
---

```{r setup, include=FALSE}
### Set knitr global options
knitr::opts_chunk$set(warning=FALSE, message=FALSE)

### Load libraries
library(data.table)
library(readr)
library(formattable)
library(DT)
library(Seurat)
```

# Purpose
To import single feature files convert to Seurat objects, and merge.  QC metrics from CellRanger are imported for each sample and exported as a single table.

## 0.1 Import QC metrics

``` {r import-metrics-files, tidy='styler', tidy.opts=list(strict=FALSE)}
samples <- list.dirs(params$crc_mtx_dir, recursive=FALSE, full.names = FALSE)
sample.dirs <- list.dirs(params$crc_mtx_dir, recursive=FALSE, full.names=TRUE)
names(sample.dirs) <- samples

metrics.dt.list <- lapply(sample.dirs, function(sd) data.table::fread(file=paste0(eval(sd), "/metrics_summary.csv")))
metrics.dt.list <- lapply(metrics.dt.list, function(dt) setnames(dt, names(dt), gsub(" ", "_", names(dt))))
metrics.dt <- rbindlist(metrics.dt.list, idcol="Sample")
sn <- metrics.dt$Sample

metrics.dt <- metrics.dt[, lapply(.SD, function(c) gsub("%|,", "", c))]
cols_convert <- names(metrics.dt)[2:ncol(metrics.dt)]
metrics.dt <- metrics.dt[, lapply(.SD, as.integer), .SDcols=cols_convert]
metrics.dt$SampleID <- sn

setcolorder(metrics.dt, "SampleID")

dir.create("./../processedData/", recursive=TRUE, showWarnings = FALSE)
data.table::fwrite(metrics.dt, file="./../processedData/cellranger_metrics.txt", sep="\t")

```

## 0.2 Create sample level rds file, if starting from filtered_matrix_bc file for each sample

```{r create-sample-seurat, tidy='styler', tidy.opts=list(strict=FALSE)}

# Convenience function for creating seurat object files
create_seurat <- function(sample_dir) {
    expression_matrix <- Seurat::ReadMtx(mtx = paste0(eval(sample_dir), "/filtered_feature_bc_matrix/matrix.mtx.gz"),
                                 features = paste0(eval(sample_dir), "/filtered_feature_bc_matrix/features.tsv.gz"),
                                 cells = paste0(eval(sample_dir), "/filtered_feature_bc_matrix/barcodes.tsv.gz"),
                                 feature.column = 2,
                                 mtx.transpose=FALSE)
    seurat <- CreateSeuratObject(expression_matrix, min.features=100)
    return(seurat)
}


seurat.list <- lapply(sample.dirs, function(sd) create_seurat(sd))

```


## 0.4 Merge sample level rds files into experiment level rds file

```{r merge-seurat, tidy='styler', tidy.opts=list(strict=FALSE), eval=params$crc_merge_sample_seurat}

num_seurat_obj <- length(seurat.list)
experiment_seurat <- merge(x=seurat.list[[1]], 
                           y=seurat.list[2:num_seurat_obj],  
                           add.cell.ids=names(seurat.list))
meta <- experiment_seurat@meta.data
experiment_seurat[["SampleID"]] <- unlist(tstrsplit(rownames(meta), "_", keep=1))
saveRDS(experiment_seurat, file=params$raw_seurat)

```

# Information about the Session 

<details>
<summary>Information about R, the OS and attached or loaded packages</summary>
```{r sesion_info}
pander::pander(sessionInfo(), compact = FALSE)
```
</details>