---
title: "Consensus clustering workup"
author: "Kim Roche"
date: "9/13/2022"
output: html_document
---

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  fig.width = 5,
  fig.height = 4
)

library(Seurat)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(ggridges)
library(caret)
library(cowplot)
library(tidyr)
library(magrittr)
library(stringr)
library(LaCroixColoR)
library(kableExtra)
```

We're going to re-run the clustering on Chang's data *within* immune cell types.

```{r}
# The Harmony-adjusted but unclustered data set seems to have been deleted.
# seurat <- readRDS("harmony/seurat_27May2022_0418UTC_10k_harmony.rds")

# The subclusterings of MoMac and DC cells have already be done in:
#   2022_06_03_Subclustering_DC.Rmd
#   2022_06_06_Subclustering_Tcells.Rmd
#   2022_06_16_Subclustering_MoMac.Rmd

momac_seurat <- readRDS("~/rosa_hernansaiz/Chang_momacdc/seurat_16Jun22_momac.rds")
dc_seurat <- readRDS("~/rosa_hernansaiz/Chang_momacdc/seurat_3Jun22_dc.rds")
tcells_seurat <- readRDS("~/rosa_hernansaiz/Chang_momacdc/seurat_6Jun22_tcells.rds")

res_lists <- list(momac = c(0.1, 0.2, 0.3, 0.5, 0.8, 1.1, 1.4),
                  dc = c(0.1, 0.2, 0.3, 0.5, 0.8, 1.1, 1.4),
                  tcells = c(0.075, 0.1, 0.15, 0.2, 0.3, 0.35, 0.4, 0.5, 0.8))

df<-data.frame("cluster"=momac_seurat@meta.data$RNA_snn_res.0.3,
               "UMAP1"=momac_seurat@reductions$umap@cell.embeddings[,1],
               "UMAP2"=momac_seurat@reductions$umap@cell.embeddings[,2]
               )

plot_all_clusterings <- function(df) {
  for(i in 1:length(res_list)) {
    coords <- df@reductions$umap@cell.embeddings
    clusters <- df@meta.data[[paste0("RNA_snn_res.", res_list[i])]]
    K <- length(unique(clusters))
    title <- paste0(K, "-cluster assignment (resolution = ", res_list[i], ")")
    
    show(
      
      ggplot(df,
                aes(x = UMAP1, y = UMAP2, fill = cluster)) +
           geom_point(size = 2, shape = 21) +
           theme_sara() +
#           scale_fill_manual(values = lacroix_palette("PeachPear", n = K, type = "continuous")) +
           labs(x = "UMAP 1", y = "UMAP 2", title = title, fill = "Cluster")
      
      )
    
  }
}

get_modularities <- function(df, res_list) {
  iteration_df <- data.frame(res = res_list,
                             max_modularity = NA)
  for(i in 1:nrow(iteration_df)) {
    txt <- capture.output(df2 <- Seurat::FindClusters(df, algorithm=2, 
                                                      resolution=iteration_df$res[i],
                                                      verbose=TRUE))
    iteration_df$max_modularity[i] <- as.numeric(str_match(txt[7], ".*?: (.*?)$")[1,2])
  }
  iteration_df
}

theme_sara <- function(){
  theme_bw(base_size=18)+
    theme(axis.text=element_text(color="black"),
          panel.background=element_rect(color="white"),
          strip.text = element_text(size=12),
          strip.background = element_rect(fill="white"))
}
```

Let's visualize the various resolutions here.

```{r, fig.height = 4, fig.width = 4.5}
plot_all_clusterings(momac_seurat, res_lists[["momac"]])
res_list<-res_lists[["momac"]]

for(i in 1:length(res_list)) {
    coords <- momac_seurat@reductions$umap@cell.embeddings
    clusters <- momac_seurat@meta.data[[paste0("RNA_snn_res.", res_list[i])]]
    K <- length(unique(clusters))
    title <- paste0(K, "-cluster assignment (resolution = ", res_list[i], ")")
    show(ggplot(data.frame(x = coords[,1],
                           y = coords[,2],
                           label = clusters),
                aes(x = x, y = y, fill = label)) +
           geom_point(size = 2, shape = 21) +
           theme_sara() +
           scale_fill_manual(values = lacroix_palette("PeachPear", n = K, type = "continuous")) +
           labs(x = "UMAP 1", y = "UMAP 2", title = title, fill = "Cluster"))
  }

```

```{r, fig.height = 4, fig.width = 4.5}
plot_all_clusterings(dc_seurat, res_lists[["dc"]])
```

```{r, fig.height = 4, fig.width = 4.5}
plot_all_clusterings(tcells_seurat, res_lists[["tcells"]])
```

Re-run these clusterings to pull modularity scores for each.

```{r}
get_modularities(momac_seurat, res_lists[["momac"]])
```

```{r}
get_modularities(dc_seurat, res_lists[["dc"]])
```

```{r}
get_modularities(tcells_seurat, res_lists[["tcells"]])
```

Check out cell quality (read counts) in cluster 3/6 of the MoMacs.

```{r, fig.height = 4, fig.width = 4}
ggplot(momac_seurat@meta.data, aes(x = RNA_snn_res.0.1, y = nCount_RNA)) +
  geom_boxplot() +
  theme_sara() +
  labs(x = "Cluster",
       y = "Total counts")

ggplot(momac_seurat@meta.data, aes(x = RNA_snn_res.0.1, y = nFeature_RNA)) +
  geom_boxplot() +
  theme_sara() +
  labs(x = "Cluster",
       y = "Total genes")
```

Finding the nearest neighbors of a randomly selected cell.

```{r}
momac_nn <- Seurat::FindNeighbors(momac_seurat, reduction="harmony", return.neighbor = TRUE)

find_knn_orig <- function(k) {
  unique_cells <- c(colnames(momac_nn)[sample(1:length(colnames(momac_nn)), size = 1)])
  counter <- 100
  while(length(unique_cells) < k & counter > 0) {
    cul_de_sac <- TopNeighbors(momac_nn[["RNA.nn"]], cell = unique_cells, n = 10)
    unique_cells <- unique(c(unique_cells, cul_de_sac))
    counter <- counter - 1
  }
  unique_cells
}

find_knn <- function(k) {
  sampled_cells <- list(sample(colnames(momac_nn), size = 1))
  counter <- 1
  while(length(unlist(sampled_cells)) < k & counter < 100) { # limit total iterations
    neighbors <- TopNeighbors(momac_nn[["RNA.nn"]], cell = sampled_cells[[counter]], n = 20)
    new_additions <- setdiff(neighbors, unlist(sampled_cells))
    # Add as many as we need
    n_cells_needed <- min(k - length(unlist(sampled_cells)), length(new_additions))
    new_additions <- sample(new_additions, size = n_cells_needed)
    sampled_cells[[counter + 1]] <- new_additions
    counter <- counter + 1
  }
  unlist(sampled_cells)
}
```

```{r, fig.height = 4, fig.width = 4}
unique_cells <- find_knn(500)

# Plot these 500
plot_df <- data.frame(x = momac_nn@reductions$umap@cell.embeddings[,1],
                  y = momac_nn@reductions$umap@cell.embeddings[,2],
                  label = colnames(momac_nn) %in% unique_cells)
ggplot() +
  geom_point(data = plot_df %>% filter(!label),
             mapping = aes(x = x, y = y),
             size = 2, shape = 21, fill = "gray") +
  geom_point(data = plot_df %>% filter(label),
             mapping = aes(x = x, y = y),
             size = 2, shape = 21, fill = "red") +
  theme_sara() +
  labs(x = "UMAP 1", y = "UMAP 2")
```

Contrast a random sampling.

```{r, fig.height = 4, fig.width = 4}
unique_cells <- sample(colnames(momac_nn), size = 1000)

# Plot these 500
plot_df <- data.frame(x = momac_nn@reductions$umap@cell.embeddings[,1],
                  y = momac_nn@reductions$umap@cell.embeddings[,2],
                  label = colnames(momac_nn) %in% unique_cells)
ggplot() +
  geom_point(data = plot_df %>% filter(!label),
             mapping = aes(x = x, y = y),
             size = 2, shape = 21, fill = "gray") +
  geom_point(data = plot_df %>% filter(label),
             mapping = aes(x = x, y = y),
             size = 2, shape = 21, fill = "red") +
  theme_sara() +
  labs(x = "UMAP 1", y = "UMAP 2")
```

Look at differences in treatment group representation across clusters.

Chi-squared test:

```{r}
md <- momac_seurat@meta.data[,c("Group", "RNA_snn_res.0.1")]
colnames(md) <- c("tx_group", "cluster")

obj <- table(md$tx_group, md$cluster)
# obj <- table(sample(md$tx_group), md$cluster) # random; not significant
obj

# There is an association between treatment group proportion and cluster
chisq.test(obj)
```

1-vs-5 model:

```{r}
# Simpler (probably bad) version

coi <- unique(md$cluster)
toi <- unique(md$tx_group)

# Summary of proportions and counts per cluster x tx_group
md %>%
  group_by(cluster, tx_group) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  select(cluster, tx_group, n) %>%
  distinct() %>%
  group_by(cluster) %>%
  mutate(n_total = sum(n)) %>%
  ungroup() %>%
  mutate(p = n/n_total) %>%
  select(cluster, tx_group, p, n) %>%
  dplyr::rename(proportion = p) -> test_df

results <- NULL
for(cluster_of_interest in coi) {
  for(tx_of_interest in toi) {
    # Non-parametric
    f <- ecdf(test_df %>%
                filter(cluster != cluster_of_interest & tx_group == tx_of_interest) %>%
                pull(proportion))

    observed_val <- test_df %>%
      filter(cluster == cluster_of_interest & tx_group == tx_of_interest) %>%
      pull(proportion)
    p <- 2 * min(f(observed_val), 1 - f(observed_val)) # two tailed
    
    results <- rbind(results,
                     data.frame(cluster = cluster_of_interest,
                                tx_group = tx_of_interest,
                                p = p))
  }
}

results$p.adj <- p.adjust(results$p, method = "BH")

results %>%
  arrange(cluster, tx_group) %>%
  kbl() %>%
  kable_styling()
```

Visualizing the composition (in terms of treatment groups) of random clusters:

```{r}
counts <- md %>%
  group_by(cluster, tx_group) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  select(cluster, tx_group, n) %>%
  distinct()

n_sample <- counts %>%
  filter(cluster == 5) %>%
  pull(n) %>%
  sum()

# Pull 20 truly random samples
sampled_clusters <- NULL
for(i in 1:20) {
  sampled_clusters <- rbind(sampled_clusters,
                            data.frame(cluster = i,
                                       tx_group = md[sample(1:nrow(md), size = n_sample),]$tx_group))
}

sampled_clusters %>%
  group_by(cluster, tx_group) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  distinct() %>%
  ggplot(aes(x = cluster, y = n, fill = tx_group)) +
  geom_bar(position = "stack", stat = "identity") +
  theme_sara() +
  labs(x = "Sampled 'cluster' (n=497)", y = "Count", fill = "Treatment group")

# Pull 40 spatially sensible clusters
sampled_clusters <- NULL
for(i in 1:40) {
  cell_ids <- find_knn(n_sample)
  mapping <- data.frame(name = cell_ids) %>%
    left_join(data.frame(name = colnames(momac_nn),
                         index = 1:ncol(momac_nn)), by = "name")
  tx_group <- momac_nn@meta.data$Group[mapping$index]
  sampled_clusters <- rbind(sampled_clusters,
                            data.frame(cluster = i,
                                       tx_group = tx_group))
}

sampled_clusters %>%
  group_by(cluster, tx_group) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  group_by(cluster) %>%
  mutate(total_n = n()) %>%
  ungroup() %>%
  mutate(proportion = n/total_n) %>%
  distinct() %>%
  ggplot(aes(x = cluster, y = proportion, fill = tx_group)) +
  geom_bar(position = "stack", stat = "identity") +
  theme_sara() +
  labs(x = "Sampled 'cluster'", y = "Count", fill = "Treatment group")
```

Better, bootstrapping null model:

```{r}
sample_randomly <- FALSE # randomly in feature space, if FALSE, sampling is of
                         # contiguous points - i.e. neighbors in the Seurat graph

if(!sample_randomly) {
  library(doParallel)
  library(foreach)
  
  doParallel::registerDoParallel(cl = parallel::makeCluster(detectCores()))
}

results <- NULL

start <- Sys.time()
for(cluster_of_interest in coi) {
  for(tx_of_interest in toi) {
    cat(paste0(cluster_of_interest, " x ", tx_of_interest, "\n"))
    
    n_sample <- md %>%
      filter(cluster == cluster_of_interest) %>%
      nrow()
    n_observed <- md %>%
      filter(cluster == cluster_of_interest & tx_group == tx_of_interest) %>%
      nrow()
    p_observed <- n_observed / n_sample
    # We need to use proportions because I'm not able to sample exactly n_sample
    # cells, I'm usually pulling slightly more. This meants comparing observed
    # vs. expected counts isn't apples-to-apples.
    
    observed <- numeric(1000)
    if(sample_randomly) {
      background <- md %>%
        filter(cluster != cluster_of_interest) %>%
        pull(tx_group)
      for(i in 1:length(observed)) {
        observed[i] <- sum(sample(background, size = n_sample, replace = T) == tx_of_interest)
      }
    } else {
      sampled_cells_to_use <- list()
      for(i in 1:length(observed)) {
        if(i %% 100 == 0) {
          cat(paste0("Random draw iteration ", i, "\n"))
        }
        sampled_cells_to_use[[i]] <- find_knn(n_sample)
      }
      cell_ids <- colnames(momac_nn)
      groups <- momac_nn@meta.data$Group
      observed <- foreach(i = 1:length(observed), .combine = 'c') %dopar% {
        library(magrittr)
        library(dplyr)
        cell_ids_use <- sampled_cells_to_use[[i]]
        mapping <- data.frame(name = cell_ids_use) %>%
          left_join(data.frame(name = cell_ids,
                               index = 1:length(cell_ids)), by = "name")
        tx_group <- groups[mapping$index]
        sum(tx_group == tx_of_interest)/length(cell_ids_use)
      }
    }
    
    # Calculate a p-value
    f <- ecdf(observed)
    p <- 2 * min(f(p_observed), 1 - f(p_observed)) # two tailed
    
    results <- rbind(results,
                     data.frame(cluster = cluster_of_interest,
                                tx_group = tx_of_interest,
                                p = p,
                                observed_p = p_observed,
                                expected_p = mean(observed),
                                CI95_lower = unname(quantile(observed, probs = c(0.025))),
                                CI95_upper = unname(quantile(observed, probs = c(0.975)))))
  }
}
results$p.adj <- p.adjust(results$p, method = "BH")
# saveRDS(results, "tx_enrichment_null_distro_summary.rds")
print(Sys.time() - start)

results %>%
  left_join(test_df %>%
              select(cluster, tx_group, n), by = c("cluster", "tx_group")) %>%
  select(cluster, tx_group, n, observed_p, expected_p, CI95_lower, CI95_upper, p, p.adj) %>%
  mutate(observed_p = round(observed_p, 3),
         expected_p = round(expected_p, 3),
         CI95_lower = round(CI95_lower, 3),
         CI95_upper = round(CI95_upper, 3)) %>%
  arrange(cluster, tx_group) %>%
  kbl() %>%
  kable_styling()
```

```{r, fig.height = 3, fig.width = 3}
# Visualize the null distribution
ggplot(data.frame(x = observed), aes(x = x)) +
  geom_histogram(color = "white") +
  geom_vline(xintercept = p_observed, color = "red") +
  theme_sara() +
  labs(x = paste0(tx_of_interest, " fraction in sampled cluster"))
```


Visualizing the observed data (as unlabeled proportions):

```{r}
ggplot(test_df, aes(x = cluster, y = proportion, fill = tx_group)) +
  geom_bar(position = "stack", stat = "identity") +
  theme_sara() +
  labs(x = "Cluster", y = "Proportion", fill = "Treatment group")
```

As unlabeled counts:

```{r}
# Counts
ggplot(test_df, aes(x = cluster, y = n, fill = tx_group)) +
  geom_bar(position = "stack", stat = "identity") +
  theme_sara() +
  labs(x = "Cluster", y = "Count", fill = "Treatment group")
```

As labeled proportions:

```{r, fig.height = 3.5, fig.width = 4.25}
# Labeled proportions
label_df <- test_df %>%
  left_join(test_df %>%
              select(-n) %>%
              pivot_wider(id_cols = c(cluster), names_from = tx_group, values_from = proportion),
            by = c("cluster")) %>%
  mutate(
    y = case_when(
      tx_group == "RT_plus_CpG" ~ RT_plus_CpG/2,
      tx_group == "RT" ~ RT_plus_CpG + RT/2,
      tx_group == "CpG" ~ RT_plus_CpG + RT + CpG/2,
      TRUE ~ RT_plus_CpG + RT + CpG + Control/2
    )) %>%
  select(-c(RT, RT_plus_CpG, CpG, Control))

ggplot(test_df, aes(x = cluster, y = proportion, fill = tx_group)) +
  geom_bar(position = "stack", stat = "identity") +
  geom_label(data = label_df,
             mapping = aes(x = cluster, y = y, label = n),
             fill = "white") +
  theme_sara() +
  labs(x = "Cluster", y = "Proportion", fill = "Treatment group")
```












