---
title: "December 2022 revisions"
author: "Kim Roche"
date: "12/11/2022"
output: html_document
---

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  fig.width = 5,
  fig.height = 4
)

library(Seurat)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(tidyr)
library(magrittr)
library(stringr)
library(kableExtra)
library(ProjecTILs)
library(LaCroixColoR)
library(cowplot)
```

Load data and ggplot2 themes.

```{r}
momac_seurat <- readRDS("~/vincent_perez/data/seurat_16Jun22_momac.rds")
dc_seurat <- readRDS("~/vincent_perez/data/seurat_3Jun22_dc.rds")
seurat_tcells <- readRDS("~/vincent_perez/data/seurat_6Jun22_tcells.rds")

theme_sara <- function(){
  theme_bw(base_size=10)+
    theme(axis.text=element_text(color="black"),
          panel.background=element_rect(color="white"),
          strip.text = element_text(size=12),
          strip.background = element_rect(fill="white"))
}

theme_sara_90 <- function() {
  theme_bw(base_size=18)+
    theme(axis.text.x=element_text(angle=90,hjust = 1,vjust = 0.5),
          axis.text=element_text(color="black"),
          panel.background=element_rect(color="black"),
          strip.text = element_text(size=12),
          strip.background = element_rect(fill="white"))
}
```

Visualize the subclustering coarsest resolution subclustering (0.35) which 
splits former cluster 1.

```{r, fig.height = 3, fig.width = 4}
# Dummy plot; save the legend from this one
p <- ggplot(data.frame(x = seurat_tcells@reductions$umap@cell.embeddings[,1],
                  y = seurat_tcells@reductions$umap@cell.embeddings[,2],
                  cluster = seurat_tcells@meta.data$RNA_snn_res.0.35),
       aes(x = x, y = y, fill = cluster)) +
  geom_point(size = 4, shape = 21) +
  theme_sara() +
  scale_fill_manual(values = lacroix_palette(type = "paired")) +
  labs(fill = "Cluster")

legend <- get_legend(p)

p <- ggplot(data.frame(x = seurat_tcells@reductions$umap@cell.embeddings[,1],
                  y = seurat_tcells@reductions$umap@cell.embeddings[,2],
                  cluster = seurat_tcells@meta.data$RNA_snn_res.0.35),
       aes(x = x, y = y, fill = cluster)) +
  geom_point(size = 2, shape = 21) +
  theme_sara() +
  scale_fill_manual(values = lacroix_palette(type = "paired")) +
  labs(x = "UMAP 1", y = "UMAP 2") +
  theme(legend.position = "none")

plot_grid(p, legend, ncol = 2, rel_widths = c(1,0.25))
```

Recreate the T-cell subtype prediction following Rosa's code in 
`2022_05_31_Subclustering_MoMacDC_Harmony.Rmd`. This uses a package called
ProjecTILs, available on Github.

```{r}
ref <- ProjecTILs::load.reference.map()

# tcell_filt.projected <- ProjecTILs::make.projection(seurat_tcells, ref=ref, filter.cells=TRUE)
# tcell_filt.projected <- ProjecTILs::cellstate.predict(ref=ref, query=tcell_filt.projected)
# 
# mapping <- data.frame(sample_name = rownames(seurat_tcells@meta.data),
#                       index = 1:nrow(seurat_tcells@meta.data)) %>%
#   left_join(data.frame(sample_name = names(tcell_filt.projected$functional.cluster),
#                        value = unname(tcell_filt.projected$functional.cluster)),
#             by = "sample_name")
# mapping$value[is.na(mapping$value)] <- "Other"
# seurat_tcells@meta.data$functional.cluster <- mapping$value

### Write to disk
# readr::write_rds(seurat_tcells, file="~/vincent_perez/data/seurat_12_11_tcells_projectils.rds")
seurat_tcells<-readRDS(file="~/vincent_perez/data/seurat_12_11_tcells_projectils.rds")
```

Visualize the ProjecTILs labels on the UMAP embedding.

```{r, fig.height = 3, fig.width = 4.25}
# Dummy plot; save the legend from this one
p <- ggplot(data.frame(x = seurat_tcells@reductions$umap@cell.embeddings[,1],
                  y = seurat_tcells@reductions$umap@cell.embeddings[,2],
                  cluster = seurat_tcells@meta.data$functional.cluster),
       aes(x = x, y = y, fill = cluster)) +
  geom_point(size = 4, shape = 21) +
  theme_sara() +
  scale_fill_manual(values = lacroix_palette(type = "paired")) +
  labs(fill = "Cluster")

legend <- get_legend(p)

p <- ggplot(data.frame(x = seurat_tcells@reductions$umap@cell.embeddings[,1],
                  y = seurat_tcells@reductions$umap@cell.embeddings[,2],
                  cluster = seurat_tcells@meta.data$functional.cluster),
       aes(x = x, y = y, fill = cluster)) +
  geom_point(size = 2, shape = 21) +
  theme_sara() +
  scale_fill_manual(values = lacroix_palette(type = "paired")) +
  labs(x = "UMAP 1", y = "UMAP 2") +
  theme(legend.position = "none")

plot_grid(p, legend, ncol = 2, rel_widths = c(1,0.25))
```

A couple of additional visuals, splitting out these classes:
* Activated CD8+ ("CD8_EarlyActiv" and "CD8_EffectorMemory")
* Exhausted CD8+ ("CD8_Tex")
* Treg ("Treg")
* Th1 ("Th1")
* Th2 ("Tfh" ???)

Visualize the ProjecTILs labels on the UMAP embedding.

```{r, fig.height = 4, fig.width = 4.5}
plot_df <- data.frame(x = seurat_tcells@reductions$umap@cell.embeddings[,1],
                      y = seurat_tcells@reductions$umap@cell.embeddings[,2],
                      flag = seurat_tcells@meta.data$functional.cluster %in% c("Tfh"))
ggplot() +
  geom_point(data = plot_df %>% filter(!flag),
             mapping = aes(x = x, y = y),
             size = 2, shape = 21, fill = "gray") +
  geom_point(data = plot_df %>% filter(flag),
             mapping = aes(x = x, y = y),
             size = 2, shape = 21, fill = "red") +
  theme_sara() +
  labs(x = "UMAP 1", y = "UMAP 2", title = "Tfh (follicular helper)")
```

Let's look at how Seurat clusters and ProjecTILs labels overlap.

```{r}
table(seurat_tcells@meta.data$RNA_snn_res.0.35,
      seurat_tcells@meta.data$functional.cluster)
```

Re-run Seurat's `FindAllMarkers()` on these *new*, finer resolution T-cell 
clusters.

```{r}
filename <- "~/vincent_perez/data/de_tcells_res0.35_vp.csv"
if(file.exists(filename)) {
  markers.list <- read.table(filename, sep = ",", header = T)
} else {
  Idents(seurat_tcells) <- seurat_tcells@meta.data$RNA_snn_res.0.35
  markers.list <- FindAllMarkers(seurat_tcells, 
                                 only.pos = TRUE,
                                 min.pct = 0.25, 
                                 logfc.threshold = .1,
                                 random.seed = 888)
  
  # Write these results out to csv files
  # markers.list %<>%
  #   filter(p_val_adj < 0.05) %>%
  #   select(cluster, gene, p_val, avg_log2FC, p_val_adj) %>%
  #   arrange(cluster, desc(avg_log2FC))

  markers.list %>%
  write.table(file = filename, quote = T, sep = ",", row.names = F)
}
```

Visualize the top markers, per-cluster, in a ComplexHeatmap.

```{r, fig.height=12, fig.width=12}
# Optionally exclude cluster 12, with its outlier expression
include_clusters <- 0:11
# include_clusters <- 0:12

# counts <- seurat_tcells@assays$RNA@counts

# Use scaled counts instead
obj_scaled <- ScaleData(seurat_tcells, features = rownames(seurat_tcells@assays$RNA))
counts <- obj_scaled@assays$RNA@scale.data

# counts2 <- counts
counts2 <- counts[,seurat_tcells@meta.data$RNA_snn_res.0.35 %in% include_clusters]

sorted_results_top10 <- markers.list %>%
  filter(p_val_adj < 0.05) %>%
  mutate(log2fc = avg_log2FC) %>%
  group_by(cluster) %>%
#  arrange(desc(log2fc)) %>%
  filter(row_number() <= 10) %>%
  ungroup() %>%
  arrange(cluster) %>% #, desc(log2fc)) %>%
  filter(cluster %in% include_clusters)

# Summarize per-cluster expression for each gene of interest
counts_subset <- counts2[rownames(counts2) %in% sorted_results_top10$gene,]
counts_long <- counts_subset %>%
  as.matrix() %>%
  as.data.frame() %>%
  cbind(gene = rownames(counts_subset)) %>%
  pivot_longer(!gene, names_to = "sample_id", values_to = "counts") %>%
  left_join(data.frame(sample_id = rownames(seurat_tcells@meta.data),
                       cluster = seurat_tcells@meta.data$RNA_snn_res.0.35),
            by = "sample_id") %>%
  group_by(cluster, gene) %>%
  summarize(mean_counts = mean(counts))

counts_wide <- counts_long %>%
  pivot_wider(id_cols = gene, names_from = "cluster", values_from = "mean_counts")

# Final wrangling to get a (6 x 10) x 
content_mat <- sorted_results_top10 %>%
  select(cluster, gene) %>%
  arrange(cluster, gene) %>%
  left_join(counts_wide, by = "gene")
row_labels <- paste0(unname(unlist(content_mat[,2])),
                     " (cl. ",
                     unname(unlist(content_mat[,1])),
                     ")")
row_labels_clusters <- paste0("cl. ", unname(unlist(content_mat[,1])))
content_mat <- as.matrix(content_mat[,3:ncol(content_mat)])
rownames(content_mat) <- row_labels
# content_mat <- log2(content_mat + 1)
column_labels <- paste("cl.", unique(sorted_results_top10$cluster))

# cpalette_base <- brewer.pal(ncol(content_mat), "Dark2")
cpalette_base <- lacroix_palette(type = "paired")[1:ncol(content_mat)]
names(cpalette_base) <- column_labels

ra <- ComplexHeatmap::rowAnnotation(`Cluster` = row_labels_clusters,
                                    col = list(`Cluster` = cpalette_base),
                                    show_legend = FALSE)

ta <- ComplexHeatmap::HeatmapAnnotation(`Cluster` = column_labels,
                                        col = list(`Cluster` = cpalette_base),
                                        show_legend = FALSE)

grey_blue_yel2 <- colorRampPalette(c("grey17", "#31446B", "#666870", "#d3c064", "#FFE945"))(256)

ht <- ComplexHeatmap::Heatmap(content_mat,
                              name = "Normalized expression",  
                              cluster_columns = FALSE,
                              cluster_rows = FALSE,
                              column_split = column_labels,
                              row_split = row_labels_clusters,
                              row_names_gp = grid::gpar(fontsize = 11),
                              column_title = character(0),
                              column_gap = unit(0.5, "mm"),
                              col = grey_blue_yel2,
                              top_annotation = ta,
                              show_column_names = FALSE,
                              row_title_rot = 0,
                              left_annotation = ra,
                              row_names_max_width = unit(6, "in"))

# filename <- "ht_deg_tcells_res0.35_kim.png"
# png(filename, width = 18, height = 18, units = "in", res = 300)
ComplexHeatmap::draw(ht, merge_legend = TRUE)
# dev.off()

pdf(file="~/vincent_perez/figures/t_cell_clusters.pdf", width = 18, height = 18)
ComplexHeatmap::draw(ht, merge_legend = TRUE)
dev.off()

```

Let's look at the expression of cluster 1's not particularly strongly expressed
or distinct markers as Seurat FeaturePlots.

```{r, fig.height = 4, fig.width = 4.5}
# Expression in raw counts
FeaturePlot(seurat_tcells, features=c("Ccl4"), combine=FALSE)
FeaturePlot(seurat_tcells, features=c("Ccl5"), combine=FALSE)
FeaturePlot(seurat_tcells, features=c("Cd3d"), combine=FALSE)
FeaturePlot(seurat_tcells, features=c("Cd3g"), combine=FALSE)
FeaturePlot(seurat_tcells, features=c("Cd8a"), combine=FALSE)
FeaturePlot(seurat_tcells, features=c("Cd8b1"), combine=FALSE)
FeaturePlot(seurat_tcells, features=c("Cxcr6"), combine=FALSE)
FeaturePlot(seurat_tcells, features=c("Ly6c2"), combine=FALSE)
FeaturePlot(seurat_tcells, features=c("Pdcd1"), combine=FALSE)
FeaturePlot(seurat_tcells, features=c("S100a4"), combine=FALSE)

# Super high, localized expression of HLA genes in tiny cluster 12
# FeaturePlot(seurat_tcells, features=c("H2-Aa"), combine=FALSE)
```

Calculate the M2 module enrichment scores for the *new*, finer resolution T-cell
clusters.

```{r}
# Tau function from Vincent
# Note this cannot handle any NAs
fTau <- function(x) {
  if(all(!is.na(x))) {
    if(min(x, na.rm=TRUE) >= 0) {
      if(max(x)!=0){
        x <- (1-(x/max(x)))
        res <- sum(x, na.rm=TRUE)
        res <- res/(length(x)-1)
      } else {
        res <- 0
      }
    } else {
      res <- NA
    } 
  } else {
    res <- NA
  } 
  return(res)
}

filename <- "~/vincent_perez/data/fsgsea_tcells_res0.35_kim.csv"
cpm <- RelativeCounts(seurat_tcells@assays$RNA@counts, scale.factor = 1e6, verbose = TRUE)
coi <- unique(seurat_tcells@meta.data$RNA_snn_res.0.35)
if(file.exists(filename)) {
  df <- read.table(filename, sep = ",", header = T)
} else {
  # ------------------------------------------------------------------------------
  #   Calculate enrichment scores (via FGSEA) on CPM
  # ------------------------------------------------------------------------------
  
  # Get pathways of interest
  gene_modules.list <- list(M2 = fgsea::gmtPathways("~/vincent_perez/data/m2.reactome.gmt"))
  
  results <- NULL
  for(cl in coi) {
    cat(paste0("Cluster ", cl, " vs. others...\n"))
    # Cluster A vs. others
    # This could be parallelized but the whole rest of this chunk takes ~1 min. 20 
    # seconds to run
    meanA <- cpm[,seurat_tcells@meta.data$RNA_snn_res.0.35 == cl]
    meanA <- rowMeans(meanA)
    meanB <- cpm[,seurat_tcells@meta.data$RNA_snn_res.0.35 != cl]
    meanB <- rowMeans(meanB)
    
    # Eliminate any genes with really low average CPM
    # This is 2/3 of genes (!)
    elim <- meanA < 1 | meanB < 1
    meanA <- meanA[!elim]
    meanB <- meanB[!elim]
    
    log2fc <- log2(meanA / meanB)
    
    # Need to filter out modules where insufficient numbers of genes are present
    # (at more than negligible abundance) in the CPM expression matrix
    use_modules <- logical(length(gene_modules.list[[1]]))
    for(i in 1:length(gene_modules.list[[1]])) {
      if(sum(unlist(gene_modules.list[[1]][i]) %in% names(meanA)) >= 1) {
        use_modules[i] <- TRUE
      }
    }
    
    # Try fgsea
    res <- fgsea(gene_modules.list[[1]][use_modules], log2fc)
    results <- rbind(results,
                     cbind(cluster = as.character(cl),
                           res[,1:7]))
  }
  
  results$cluster <- factor(results$cluster, levels = 0:12)

  # ------------------------------------------------------------------------------
  #   Complex normalized enrichment scores (NES) with specificity (tau)
  # ------------------------------------------------------------------------------
  
  min_NES <- min(results$NES, na.rm = T)
  
  results_tau <- results %>%
    filter(!is.na(NES)) %>%
    mutate(NES_positive = NES + abs(min_NES)) %>%
    group_by(pathway) %>%
    mutate(tau = fTau(NES_positive)) %>%
    ungroup()
  
  # ------------------------------------------------------------------------------
  #   Write out full results
  # ------------------------------------------------------------------------------
  
  # All results
  sorted_results <- results_tau %>%
    group_by(cluster) %>%
    # arrange(desc(NES)) %>%
    arrange(desc(NES*tau)) %>%
    ungroup() %>%
    # arrange(cluster, desc(NES)) %>%
    arrange(cluster, desc(NES*tau))
  
  # write.table(sorted_results, file = filename, quote = T, sep = ",", row.names = F)
}
```

Visualize the top 10 pathways, per-cluster, in a ComplexHeatmap.

```{r, fig.height=12, fig.width=12}
# ------------------------------------------------------------------------------
#   Pull top 10 for heatmap
# ------------------------------------------------------------------------------

sorted_results_all <- read.table(filename, sep = ",", header = T)
sorted_results_top10 <- sorted_results_all %>%
  group_by(cluster) %>%
  arrange(desc(NES*tau)) %>%
  filter(row_number() <= 10) %>%
  ungroup() %>%
  arrange(cluster, desc(NES*tau))

# Build a table we'll write out to a .csv
map <- data.frame(index = 1:nrow(sorted_results_top10),
                  pathway = sorted_results_top10$pathway)

content <- map %>%
  left_join(sorted_results_all %>%
              select(pathway, cluster, NES), by = "pathway") %>%
  pivot_wider(id_cols = c(index, pathway), names_from = cluster, values_from = NES, names_sort = T)

content_mat <- as.matrix(content[,3:ncol(content)])
rownames(content_mat) <- sorted_results_top10$pathway

df <- cbind(Cluster = rep(0:(ncol(content_mat)-1), each = 10),
            `Top 10 gene sets` = rownames(content_mat),
            as.data.frame(content_mat))
rownames(df) <- NULL
colnames(df)[3:ncol(df)] <- paste0("Cluster ", colnames(df)[3:ncol(df)], " NES")

row_labels <- paste("cl.", unname(sorted_results_top10$cluster))
column_labels <- paste("cl.", unique(sorted_results_top10$cluster))

# cpalette_base <- brewer.pal(ncol(content_mat), "Dark2")
cpalette_base <- lacroix_palette(type = "paired")[1:ncol(content_mat)]
names(cpalette_base) <- column_labels

ra <- ComplexHeatmap::rowAnnotation(`Cluster` = row_labels,
                                    col = list(`Cluster` = cpalette_base),
                                    show_legend = FALSE)

ta <- ComplexHeatmap::HeatmapAnnotation(`Cluster` = column_labels,
                                        col = list(`Cluster` = cpalette_base),
                                        show_legend = FALSE)

grey_blue_yel2 <- colorRampPalette(c("grey17", "#31446B", "#666870", "#d3c064", "#FFE945"))(256)

ht <- ComplexHeatmap::Heatmap(content_mat,
                              name = "NES",  
                              cluster_columns = FALSE,
                              cluster_rows = FALSE,
                              column_split = column_labels,
                              row_split = row_labels,
                              row_names_gp = grid::gpar(fontsize = 12),
                              column_title = character(0),
                              column_gap = unit(0.5, "mm"),
                              col = grey_blue_yel2,
                              top_annotation = ta,
                              show_column_names = FALSE,
                              row_title_rot = 0,
                              left_annotation = ra,
                              row_names_max_width = unit(6, "in"))
# filename <- paste0("ht_m2_tcells_res", pieces[[1]][2], ".", pieces[[1]][3], "_kim.png")
# png(filename, width = 22, height = 19, units = "in", res = 300)
ComplexHeatmap::draw(ht, merge_legend = TRUE)
# heatmap_legend_side = "bottom"
# annotation_legend_side = "bottom"
# dev.off()
```

Render barplots of G1/S/G2M phase proportions in each cluster for T-cells.

```{r, fig.height = 3, fig.width = 6}

# T-cells
plot_df <- data.frame(cluster = seurat_tcells@meta.data$RNA_snn_res.0.35,
                      phase = seurat_tcells@meta.data$Phase)

# MoMacs
# plot_df <- data.frame(cluster = momac_seurat@meta.data$RNA_snn_res.0.1,
#                       phase = momac_seurat@meta.data$Phase)

plot_df %>%
  group_by(cluster) %>%
  mutate(n_total = n()) %>%
  ungroup() %>%
  group_by(cluster, phase) %>%
  mutate(n = n(),
         p = n/n_total) %>%
  select(cluster, phase, n, p) %>%
  distinct() %>%
  mutate(phase = factor(phase, levels = c("G1", "S", "G2M"))) %>%
  ggplot(aes(x = as.factor(cluster), y = p, fill = phase)) +
  geom_bar(stat = "identity", position = "stack") +
  theme_sara() +
  labs(x = "Cluster", y = "Proportion", fill = "Phase")
```

Look for Chang's marker genes.

```{r, fig.height = 4, fig.width = 4.5}
# Markers for different T cell populations:
# Activated CD8+ cells:
# CD3+ CD8+ Tim3+ Lag3+ PD1+ TCF1+ CD44+ CD62L- Granzyme B+ IFNg+ TNFa+ IL2+
#
# Exhausted CD8 T cells:
# CD3+ CD8+ Tim3+ Lag3+ PD1+ CTLA4+ BTLA+ TIGIT+ Eomes+ CD69+ CD44+ CD62L- TCF1- Granzyme B- IFNg- IL2- TNFa-
# 
# T regulatory cells:
# CD3+ CD4+ CD25+ Foxp3+
# 
# Th1 cells:
# CD3+ CD4+ CCR5+ CXCR3+ 
# 
# Th2 cells:
# CD3+ CD4+ CXCR4+ CCR3+ CCR4+ CCR8+ 

gene_names <- c("Cd3d", "Cd3e", "Cd3g",
                "Cd8a", "Cd8b1",
                "Havcr2",
                "Lag3",
                "Pdcd1",
                "Tcf7",
                "Cd44",
                "Sell",
                "Gzmb",
                "Ifng",
                "Il2",
                "Tnf")

gene_names <- c("Ctla4",
                "Btla",
                "Tigit",
                "Eomes",
                "Cd69")

gene_names <- c("Cd4",
                "Il2ra",
                "Foxp3")

gene_names <- c("Ccr5",
                "Cxcr3")

gene_names <- c("Cxcr4",
                "Ccr3",
                "Ccr4",
                "Ccr8")

# Search for gene names like ...
# unlist(sapply(rownames(seurat_tcells@assays$RNA@counts), function(x) {
#   grep("^Ccr8", x)
# }))

# Pull indices to check the raw counts
# idx <- sapply(gene_names, function(x) {
#   which(rownames(seurat_tcells@assays$RNA@counts) == x)
# })
# hist(unlist(seurat_tcells@assays$RNA@counts[15905,]))

for(gene in gene_names) {
  show(FeaturePlot(seurat_tcells, features=c(gene), combine=FALSE))
}
```


